---
title: "AngieRuiz_A1"
author: "Angie Ruiz"
date: "02/19/21"
output: html_document
editor_options: 
chunk_output_type: console
---


```{r}
knitr::opts_chunk$set(warning = F, message = F)
library(tidyverse)
library(plotly)
library(sf)
library(tigris)
library(leaflet)
library(censusapi)

Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")
```


First, you'll need to download all the individual zip files yourself and unzip the CSVs inside them into your working directory (in my case they're in a folder called pge). From there, you should adapt the loop code from 1.6 to be applied to more files. One way to approach this is using three nested for loops. This loop structure essentially helps you construct the different file names. Otherwise, there are two other special issues to address before the code will run smoothly:

- The year 2021 doesn't have all 4 quarters of data, so when your for loop hits 2021 Q3, it will try to read a file that doesn't exist and cause an error, which causes you to lose your whole process. You want to be able to skip files that don't exist. The `next` operation moves on to the next iteration of the loop, and would be useful here.
- The electricity and gas CSVs are similar except for the units of the energy, which means different field names. Electricity has `TOTALKWH` and `AVERAGEKWH`, and Gas has `TOTALTHM` and `AVERAGETHM`. You need to use mutate to convert these values to `TOTALKBTU` (Average KBTU isn't necessary for this assignment) and remove the original fields before rbinding the dataframes together. Since you will know whether you're dealing with an Electricity or Gas dataset using the `type` variable, you can use this within if statements, or do the equivalent inside of a pipeline.


```{r}

years <- 2017:2021
quarters <- 1:4
type1e <- "Electric"
type2g <- "Gas"

pge_elec <- NULL
pge_gas <- NULL

#Set Up for Electric Data
for(year in years) {
  for(quarter in quarters) {
      
      filename1 <- 
        paste0(
      "PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type1e,
      "UsageByZip.csv"
        )
      
  if (!file.exists(filename1))next
      
  temp <- read_csv(filename1)
  
        if(type1e == "Electric") {
        temp <-
          temp %>%
          mutate(TOTALKBTU = TOTALKWH * 3412.14) %>%
          select(-TOTALKWH, -AVERAGEKWH)
        }
  }}
  
  pge_elec <- rbind(pge_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_elec, "pge_elec.rds")
  
  
  #Set Up for Gas Data
for(year in years) {
  for(quarter in quarters) {
      
      filename2 <- 
        paste0(
      "PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type2g,
      "UsageByZip.csv"
        )
      
  if (!file.exists(filename2))next
      
  temp <- read_csv(filename2)
  
        if(type2g == "Gas") {
        temp <-
          temp %>%
          mutate(TOTALKBTU = TOTALTHM * 100000.39) %>%
          select(-TOTALTHM, -AVERAGETHM)
        }
  
  pge_gas <- rbind(pge_gas,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_gas, "pge_gas.rds")


  }
}
```

There is one genuine problem with the data that needs to be solved, but it is hard to notice at this stage, and hard to notice at all until you make your first plots. You won't be penalized for not correcting it, but be on the lookout for something odd in the data, and if you find it, think about how to address it in your for loop.

Next, you should explore the dataframe to notice that you only need 4 customer classes. Then, using the techniques from Chapter 1.7, you can manipulate the dataframe to give you total KBTU per month. To enable the dataframe to plot with time on the X-axis, it helps to create a `DATE` field that is an alphanumerically increasing string, or is a full Date object (as I'm guiding below). 

```{r}
#Filtering by customer class

pge_filter_elec <- 
  filter(
    pge_elec,
    CUSTOMERCLASS %in% 
      c(
        "Elec- Residential",
        "Elec- Commercial"
  )
  )%>% 
  
  pge_filter_gas <- 
  filter(
    pge_gas,
    CUSTOMERCLASS %in% 
      c(
        "Gas- Residential",
        "Gas- Commercial"
  )
  )%>% 

#Getting rid of the data columns we don't need
  pge_select_elec <-
  select(
    pge_filter_elec,
    -c(COMBINED, AVERAGEKWH)
  )%>% 

pge_select_gas <-
  select(
    pge_filter_gas,
    -c(COMBINED, AVERAGETHM)
  )%>% 

#Setting the categories I want to keep as discrete categories, this will give me
#the data for the whole PGE region for each month and year.
pge_group_elec <-
  group_by(
    pge_select_elec,
    MONTH,
    YEAR,
    CUSTOMERCLASS
  ) %>% 

  pge_group_gas <-
  group_by(
    pge_select_gas,
    MONTH,
    YEAR,
    CUSTOMERCLASS
  ) %>% 

#Creating new fields based on the data I choose to keep and manipulate with functions
  pge_summarize_elec <-
  summarize(
    pge_group_elec,
    TOTALKBTU = sum(TOTALBTU, na.rm = T),
    TOTALCUSTOMERS =sum(TOTALCUSTOMERS,na.rm = T),
    DATE = 
      as.Date(paste0(YEAR, "-", MONTH,"-", "01"))
  ) %>% 
  
  pge_summarize_gas <-
  summarize(
    pge_group_gas,
    TOTALKBTU = sum(TOTALBTU, na.rm = T),
    TOTALCUSTOMERS =sum(TOTALCUSTOMERS,na.rm = T),
    DATE = 
      as.Date(paste0(YEAR, "-", MONTH,"-", "01"))
  ) %>% 
  

#Combining (stacking) the electric and gas data

pge_final <- NULL

pge_final <- rbind(pge_summarize_elec , pge_summarize_gas)

pge_final_commercial <-
  filter(
    pge_final,
    CUSTOMERCLASS %in% 
      c(
        "Gas- Commercial",
        "Elec- Commercial"
  )
  )%>% 
  
pge_final_commercial


  pge_final_residential<-
  filter(
    pge_final,
    CUSTOMERCLASS %in% 
      c(
        "Gas- Residential",
        "Elec- Residential"
  )
  )%>% 
pge_final_residential

```

Next, write a `ggplot()` pipeline for the residential time series. 

```{r}

library(scales)
library(plotly)
library(ggplot)

pge_chart_residential <-
pge_final_residential %>% 
  ggplot() +
    geom_bar(
      aes(
      x = DATE,
      y = TOTALBTU,
      fill = CUSTOMERCLASS,
      
    ),
    stat = "identity",
    position = "stack"
  ) +
  
  labs(
    x = "Date",
    y = "kBTU",
    title = "PG&E Territory Monthly Residential Electrcity and Gas Usage",
    fill = "Energy Type"
  ) 

pge_chart_residential


pge_chart_commercial <-
pge_final_commercial %>% 
  ggplot() +
    geom_bar(
      aes(
      x = DATE,
      y = TOTALBTU,
      fill = CUSTOMERCLASS,
      
    ),
    stat = "identity",
    position = "stack"
  ) +
  
  labs(
    x = "Date",
    y = "kBTU",
    title = "PG&E Territory Monthly Commercial Electrcity and Gas Usage",
    fill = "Energy Type"
  ) 

pge_chart_commercial



#Analyze COVID19 Impact on Energy Usage (2019-2020)
#Filter by year

pge_commericial_covid <-
filter(
    pge_final_commercial,
    YEAR %in% 
      c(
        "2019, 2020"
  )
  )%>% 

pge_residential_covid <-
filter(
    pge_final_residential,
    YEAR %in% 
      c(
        "2019, 2020"
  )
  )%>% 

#Make Chart to demonstrate COVID19 Impact on Energy Usage on a Plot

pge_comm_cov_chart <-
  pge_commercial_covid %>% 
  ggplot() +
    geom_bar(
      aes(
      x = DATE,
      y = TOTALBTU,
      fill = CUSTOMERCLASS,
      
    ),
    stat = "identity",
    position = "stack"
  ) +
  
  labs(
    x = "Date",
    y = "kBTU",
    title = "PG&E Territory Monthly Commercial Electricity and Gas Usage in 2019-2020",
    fill = "Energy Type"
  ) 

pge_comm_cov_chart


pge_res_cov_chart <-
  pge_residential_covid %>% 
  ggplot() +
    geom_bar(
      aes(
      x = DATE,
      y = TOTALBTU,
      fill = CUSTOMERCLASS,
      
    ),
    stat = "identity",
    position = "stack"
  ) +
  
  labs(
    x = "Date",
    y = "kBTU",
    title = "PG&E Territory Monthly Residential Electricity and Gas Usage in 2019-2020",
    fill = "Energy Type"
  ) 

pge_res_cov_chart
  
```


#Assumptions: Although we know this is not the correct timeline, with the way I input the data it analyzes energy usage from all of 2019 to all of 2020 even though rates of contamination and whether or not people were working from home varied greatly during this time.


#I keep getting the error Error in ggplot(.) : object '____' not found for my final variables and my chart variables and cannot see a plot. Therefore, I am frustratingly unable to analyze the electricity and gas usage data. 


#However, if I were to try and answer the questions and analyze commercial and residential energy usage from 2019-2020 I would hypothesize that We would see an increase in residential gas and electricity usage since many people transitioned to work from home and therefore were in their houses for an additional 8 hours a day cooking, heating/using AC, charging their devices, etc.for these additional hours per day whereas prior to 2019 they wouldn't do these things while they were at their office. I would anticipate that commerical energy usage would be roughly similar because although many people were no longer working from the office, in my biased and personal perspective, many employers still kept the buildings running. Inclusively, some were semi- work from home where the building was still being used my employees a few days a week. I wish I could confirm or deny these observations but I cannot figure out where my ggplot is going wrong.

You should now write some similar code to visualize commercial energy.

Include commentary throughout this document, replacing my own commentary.title: "Ruiz_218X_HW1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
